Manuscript Draft\\ \vspace{2cm} What is long-distance dispersal? And a taxonomy of dispersal events \\ \textbf{Supplementary Material}

### 2016-06-26 Manuscript/MS- What is LDD.tex
Edits to Table 2 and Discussion.

\begin{equation}
					N_b= 4 \pi \sigma^2 \delta
\end{equation}

where $\delta$ is the density of adults per unit area and $\sigma$ is the variance in distance between birth and breeding sites. This formulation is often called the neighborhood size and assumes a normal distribution of distances between parents and offspring (out in a perfect circular shape from the source). Thus, changes in the variance of dispersal distance can affect $N_e$ (highly clumped populations will have reduced $N_e$). This is the basic model of "Isolation by Distance" proposed by   \citet{Wright:1943aa,Wright:1946aa}. Under this type of model, migration (gene flow) is given by the variance in dispersal, and not by the proportion of the population that is composed of migrants (denoted $m$), as is the case with island models \citep{Slatkin:1985qb}.

@@

\begin{equation}
  			N_b = 4 \pi (\sigma^2_s + \frac{t \sigma^2_p}{2}) \delta (1 + t)
\end{equation}

where $\sigma^2_s$ is the variance in seed dispersal, $\sigma^2_s$ is the variance in pollen dispersal and $\delta$ is the density of potential parents.

@@

The extent of neighborhood area in plants can be extremely variable, depending on life-history attributes such as life-span, spacing patterns, mating system, etc. Even a limited sample of available information (Table S1) highlights the fact that the size of neighborhood areas can in some cases exceed the geographic limits of local populations \citep{Nason:1998aa}. The size of neighborhood areas may encompass at least four orders of magnitude, $10{^-2}-10^2$ km in radius, and include many individuals. Therefore, reference to this "genetic/evolutionary" paradigm and reference \citep[sensu][]{Waples:2006ev} may be instrumental to understand the actual role of LDD events in shaping the structuring of genetic pools and contributing to gene dispersal.

Whenever there is a large discrepancy between population area extent and $N_b$ we might expect the frequency of $LDD_{loc}$ and $LDD_{neigh}$ differ enormously. For example, relatively small $N_b$ may rise the importance of $LDD_{loc}$ in preserving scenarios of panmixia within a local population, as most distant dispersal events will disseminate seeds outside the neighborhood of maternal plants.

@@

Empirical evaluation of differential contributions to the different forms of LDD events outlined in Table 1 requires identification of source trees as well as assignment of the dispersed propagules to specific vectors or functional groups of vectors \citep{Jordano:2007}. Recently, DNA-barcoding techniques have been developed and successfully applied to the identification of frugivore species contributing to specific seed dispersal events whose source can be identified with genetic, direct assignment techniques \citep{GonzalezVaro:2014ij}. Otherwise, visual identification can reliably assign the genotyped seeds to frugivore species groups based on specific characteristics of scats and regurgitations \citep{Jordano:2007a}.

@@

As defined in our framework (Table 1), LDD, and in particular $LDD_{ss}$ events are a particular case of extreme events \citep{Garcia:2017aa}.

\subsection*{Challenges and Promising Avenues for Research}

\emph{Acknowledgements}. I am indebted to Cristina García, José A. Godoy, Manolo Carrión, Juan Luis García-Castaño, Jesús Rodríguez and, especially, Juan Miguel Arroyo for generous help with field and laboratory work and making possible this study. I appreciate the help and advice of Cristina García and Etienne Klein during the final stages of the manuscript. The study was supported by a Junta de Andalucía Excellence Grant (RNM-5731), as well as a Severo Ochoa Excellence Award from the Ministerio de Economía y Competitividad (SEV-2012-0262). The Agencia de Medio Ambiente, Junta de Andalucía, provided generous facilities that made possible this study in the Andalusian natural parks (Sierra de Cazorla, Alcornocales) and authorized my work there.

@@

%\section*{References}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
%%% Unnumbered Literature Cited
\bibliographystyle{bes}        %Compile with bes.bst style file
\bibliography{MS_BES}

@@

%-------------------------------------------------------------------- Table 2
\pagestyle{empty}
\begin{landscape}
\begin{table}
\captionsetup{width=21cm}%.75\textwidth}
\caption{Frequencies of \textit{Prunus mahaleb} seed dispersal events for different frugivore groups according to population area limits and genetic neighborhood limits. See Fig. 2 for a graphical representation of the four scenarios.}
\vspace{0.5cm}
    \begin{tabular}{lp{3cm}p{3cm}p{3cm}p{3cm}}
    \hline \\
    Frugivore group & Within-population, within-neighborhood $SSD_{loc}$ & Within-population, long-distance $LDD_{loc}$ & Outside-population, within-neighborhood $LDD_{neigh}$ & Strict-sense long-distance $LDD_{ss}$ \\\\ \hline \\
    Small birds     & ?~           & ?~            & ?~           & ?~  \\
    \textit{Turdus} & ?~           & ?~            & ?~           & ?~  \\

    Large birds     & ?~           & ?~            & ?~           & ?~  \\

    Mammals         & ?~           & ?~            & ?~           & ?~  \\\\
\hline
    \end{tabular}
\end{table}
\end{landscape}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% MS NOTES
\begin{comment}
NOTES
% -------------- Crawford 21986 Heredity ------------------------------------
Sewal Wright's neighborhood model indicates that the area containing a panmictic unit within a continuous and uniform array of organisms can be 
estimated by $4\pir^2$ where a-2 is the parent-offspring dispersal variance measured around a zero  mean and relative to a single reference axis passing through the population.

>>>>

\begin{equation}
					N^b_e= 4 \pi \sigma \delta
\end{equation}

where $\delta$ is the density of adults per unit area and $\sigma$ is the standard deviation of the distance between birth and breeding sites. This formulation is often called the neighborhood size and assumes a normal distribution of distances between parents and offspring (out in a perfect circular shape from the source). Thus, changes in the variance of dispersal distance can affect $N^b_e$ (highly clumped populations will have reduced $N^b_e$). This is the basic model of "Isolation by Distance" proposed by   \citet{Wright:1943aa,Wright:1946aa}. Under this type of model, migration (gene flow) is given by the variance in dispersal, and not by the proportion of the population that is composed of migrants (denoted $m$), as is the case with island models \citep{Slatkin:1985qb}. With enough distance separating them, two plant individuals have a low probability of mating and can be considered members of distinct genetic populations even if they are not located in geographically distinct populations.

@@

\begin{equation}
  		N^b_e = 4 \pi (\sigma^2_s + \frac{t \sigma^2_p}{2}) \delta (1 + t)
\end{equation}

where $\sigma_s$ is the standard deviation of seed dispersal distance, $\sigma_p$ is the standard deviation of pollen dispersal distance, and $\delta$ is the density of potential parents.

@@

The extent of neighborhood area in plants can be extremely variable, depending on life-history attributes such as life-span, spacing patterns, mating system, etc. Even a limited sample of available information (Table S1) highlights the fact that the size of neighborhood areas can in some cases exceed the geographic limits of local populations \citep{Nason:1998aa}. The size of neighborhood areas may encompass at least four orders of magnitude, $10{^-2}-10^2$ km in radius, and include many individuals. Therefore, reference to this "genetic/evolutionary" paradigm and reference to the geographic boundaries \citep[sensu][]{Waples:2006ev} may be instrumental to understand the actual role of LDD events in shaping the structuring of genetic pools and contributing to gene dispersal.

Whenever there is a large discrepancy between population area extent and $N^b_e$ we might expect the frequency of $LDD_{loc}$ and $LDD_{neigh}$ differ enormously. For example, relatively small $N^b_e$ may rise the importance of $LDD_{loc}$ in preserving scenarios of panmixia within a local population, as most distant dispersal events will disseminate seeds outside the neighborhood of maternal plants.

@@

Empirical evaluation of differential contributions to the different forms of LDD events outlined in Table 1 requires identification of source trees as well as assignment of the dispersed propagules to specific vectors or functional groups of vectors \citep{Jordano:2007}. Recently, DNA-barcoding techniques have been developed and successfully applied to the identification of frugivore species contributing to specific seed dispersal events whose source can be identified with genetic, direct assignment techniques \citep{GonzalezVaro:2014ij}. Otherwise, visual identification can reliably assign the genotyped seeds to frugivore species groups based on specific characteristics of scats and regurgitations \citep{Jordano:2007}.

We inferred the frugivore groups contributing dispersal events by visually identifying scats and regurgitations in seed traps and line transects \citep[see ][ and Suppl. Mat. for additional details of methods]{Jordano:2007}. These frugivore functional groups include up to 38 bird and 4 mammal species feeding on \textit{P. mahaleb} fruits \citep{Jordano:2000ft}. Here, we differentiate four major frugivore groups: large carnivorous mammals (such as foxes, badgers, and stone martens); two species of medium-sized frugivorous birds, mistle thrushes (\textit{T. viscivorus}), and carrion crows (\textit{Corvus corone}); and a pool of small-sized frugivorous birds, including warblers, redstarts, and robins \citep{Jordano:2007}.

To a large extent, short-distance dispersal events (strict-sense, $SDD_{loc}$ events) are contributed by small- and medium-sized (\textit{Turdus}) frugivorous birds (Table 2). Given the relatively reduced $N^b_e$ area of \textit{P. mahaleb} (Suppl. Mat. Table S1), $< 1km^2$, well below the extent of the local study population \citep{Garcia:2007he,Garcia:2005fu},  we cannot estimate $LDD_{neigh}$ events (Table 2), as all LDD events outside the reference population occur, by definition, outside the $N^b_e$ area. Larger frugivores such as corvids and the pigeon \textit{Columba palumbus} contribute most LDD events, and most immigrant seeds potentially dispersed from other populations. Notably, strict-sense long-distance dispersal ($LDD_{ss}$) appears consistently associated with large-bodied frugivores (Table 2), most likely associated with a greater frequency of movements outside the local population (Fig. 4).

@@

Combining spatially-explicit references to the geographic population limits and the $N^b_e$ area extent helps avoiding some imprecision in setting distance thresholds to characterize LDD events. In addition the framework outlined in Table 1 bridges the combined demographic and genetic effects of LDD events.

As defined in our framework (Table 1), LDD, and in particular $LDD_{ss}$ events are a specific case of extreme events \citep{Garcia:2017aa} consistently associated with large-sized frugivores, yet including also medium-sized and highly efficient frugivorous bird species.

- Large-bodied frugivores and LDD.  Discuss lb frugivores not always contributing LDD events: Pratt, tapirs
- Loss of large-bodied frugivores, consequences.

\subsection*{Challenges and future avenues for research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Acknowledgements
\emph{Acknowledgements}. I am indebted to Cristina García, José A. Godoy, Manolo Carrión, Juan Luis García-Castaño, Jesús Rodríguez and, especially, Juan Miguel Arroyo for generous help with field and laboratory work and making possible this study. I appreciate the help and advice of Cristina García and Etienne Klein during the final stages of the manuscript. The study was supported by a Junta de Andalucía Excellence Grant (RNM-5731), as well as a Severo Ochoa Excellence Award from the Ministerio de Economía y Competitividad (SEV-2012-0262). The Agencia de Medio Ambiente, Junta de Andalucía, provided generous facilities that made possible this study in the Andalusian natural parks (Sierra de Cazorla, Alcornocales) and authorized my work there.

@@

%\section*{References}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
%%% Unnumbered Literature Cited
\bibliographystyle{bes}        %Compile with bes.bst style file
\bibliography{MS_BES}

@@

%-------------------------------------------------------------------- Table 2
\pagestyle{empty}
\begin{landscape}
\begin{table}
\captionsetup{width=21cm}%.75\textwidth}
\caption{Relative frequencies of \textit{Prunus mahaleb} seed dispersal events for different frugivore groups according to population area limits and genetic neighborhood limits. See Fig. 2 for a graphical representation of the four scenarios. \textit{N}= 655 seeds \citep[see Table 1 in ][]{Jordano:2007}. Given that the estimated neighborhood size is smaller than population area, $LDD_{neigh}$ would be zero.}
\vspace{0.5cm}
    \begin{tabular}{lp{3cm}p{3cm}p{3cm}p{3cm}p{3cm}}
    \hline \\
    Frugivore group & Within-population, within-neighborhood $SDD_{loc}$ & Within-population, long-distance $LDD_{loc}$ & Outside-population, within-neighborhood $LDD_{neigh}$ & Strict-sense long-distance $LDD_{ss}$ & \textit{N} seeds  \\\\ \hline \\
	Small-birds	    & 0.7842 &  0.0171 & 0.00 & 0.1986 & 292  \\
	\textit{Turdus} & 0.2370 &  0.5549 & 0.00 & 0.2081 & 173  \\
	Large-birds     & 0.0435 &  0.3913 & 0.00 & 0.5652 &  23  \\
	Mammals         & 0.0120 &  0.2455 & 0.00 & 0.7425 & 167  \\\\
\hline
    \end{tabular}
\end{table}
\end{landscape}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% MS NOTES
\begin{comment}
NOTES
% -------------- Crawford 21986 Heredity ------------------------------------
Sewall Wright's neighborhood model indicates that the area containing a panmictic unit within a continuous and uniform array of organisms can be 
estimated by $4\pir^2$ where a-2 is the parent-offspring dispersal variance measured around a zero  mean and relative to a single reference axis passing through the population.

### 2016-06-26 Manuscript/SupplMat.tex
Added Fig 4 and moved fig to Suppl Mat. Edited figure legends.
Edits to Table 2 and Discussion.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% MS Draft Journal of Ecology Special Issue, Dispersal - SUPPL MATERIAL
%% Pedro - 07 June 2016
%----------------------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Headers
\documentclass[a4paper, 12pt]{article}}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\pagestyle{plain}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\usepackage{lscape}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{longtable}
\usepackage{rotating} 
\usepackage{standalone}                                                                                    
\newcommand{\tab}{\hspace{5mm}}
\usepackage{tabularx} 
\usepackage[margin=10pt,font=small,labelfont=bf]{caption}
\usepackage[left]{lineno}
\usepackage{caption}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy}     % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{{\tiny Jordano - What is long-distance dispersal?}}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}
\usepackage{parskip}
\linespread{1.25}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title page
\begin{document}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Supplementary Material
%------------------------------------------------------------------- Table S1
\begin{landscape}
\begin{table}
  \caption{Summary of neighborhood area sizes and estimated neighborhood radius for tree species with different combinations of dispersal modes. Data from Nason et al and present study. }
    \begin{tabular}{lllccc}
     Species                & Pollinator  & Seed disperser & Density ($ha^{-1}$) & Breeding unit ($km^2$) & Radius (km) \\
    \textit{Ficus dugandii}          & Fig wasp    & Vertebrates    & 0.004          & 631.7               & 14.2        \\
    \textit{Ficus obtusifolia}       & Fig wasp    & Vertebrates    & 0.072          & 105.9               & 5.8         \\
    \textit{Prunus mahaleb}          & Bees, flies & Vertebrates    & 0.003          & 0.87                & 0.042       \\
    \textit{Frangula alnus}          & Bees, flies & Vertebrates    & 0.0004         & 0.45                & 0.013       \\
    \textit{Astrocaryum mexicanum}   & Beetle      & Vertebrates    & 1364.0         & 0.011               & 0.06        \\
    \textit{Calophyllum longifolium} & Bees        & Vertebrates    & 0.28           & 1.241               & 0.629       \\
    \textit{Platypodium elegans}     & Bees        & Wind           & 0.78           & 0.866               & 0.525       \\
    \textit{Cedrus atlantica}        & Wind        & Wind           & 61.7           & 0.151               & 0.22        \\
   \textit{Fraxinus americana}      & Wind        & Wind           & 24.7           & 0.008               & 0.05        \\
    \textit{Pseudotsuga menziesii}   & Wind        & Wind           & 25.0           & 0.078               & 0.158       \\
    \end{tabular}
\end{table}
\end{landscape}
\newpage

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% MS Draft Journal of Ecology Special Issue, Dispersal - SUPPL MATERIAL
%% Pedro - 07 June 2016
%----------------------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Headers
\documentclass[a4paper, 12pt]{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\pagestyle{plain}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\usepackage{lscape}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{longtable}
\usepackage{rotating}                                                                                     
\newcommand{\tab}{\hspace{5mm}}
\usepackage{tabularx} 
\usepackage[margin=10pt,font=small,labelfont=bf]{caption}
\usepackage[left]{lineno}
\usepackage{caption}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{empty}     % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{{\tiny Jordano - What is long-distance dispersal?}}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}
\usepackage{parskip}
\linespread{1.25}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title page
\begin{document}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Supplementary Material
%------------------------------------------------------------------- Table S1
\begin{landscape}
\begin{table}
  \caption{Summary of neighborhood area sizes and estimated neighborhood radius for tree species with different combinations of dispersal modes. Data from \citet{Nason:1998aa,Garcia:2005fu,Garcia:2007he} and present study.}
    \begin{tabular}{lllccc}
     Species                & Pollinator  & Seed disperser & Density ($ha^{-1}$) & Breeding unit ($km^2$) & Radius (km) \\
    \textit{Ficus dugandii}          & Fig wasp    & Vertebrates    & 0.004          & 631.7               & 14.2        \\
    \textit{Ficus obtusifolia}       & Fig wasp    & Vertebrates    & 0.072          & 105.9               & 5.8         \\
    \textit{Prunus mahaleb}          & Bees, flies & Vertebrates    & 0.003          & 0.87                & 0.042       \\
    \textit{Frangula alnus}          & Bees, flies & Vertebrates    & 0.0004         & 0.45                & 0.013       \\
    \textit{Astrocaryum mexicanum}   & Beetle      & Vertebrates    & 1364.0         & 0.011               & 0.06        \\
    \textit{Calophyllum longifolium} & Bees        & Vertebrates    & 0.28           & 1.241               & 0.629       \\
    \textit{Platypodium elegans}     & Bees        & Wind           & 0.78           & 0.866               & 0.525       \\
    \textit{Cedrus atlantica}        & Wind        & Wind           & 61.7           & 0.151               & 0.22        \\
   \textit{Fraxinus americana}      & Wind        & Wind           & 24.7           & 0.008               & 0.05        \\
    \textit{Pseudotsuga menziesii}   & Wind        & Wind           & 25.0           & 0.078               & 0.158       \\
    \end{tabular}
\end{table}
\end{landscape}
\newpage

@@

%------------------------------------------------------------------ Figure S2
\begin{figure}[htbp]
\centerline{\includegraphics[height=22cm]{FigS2.pdf}}
%
\caption{Differential contributions of functional groups of frugivores to the short- ($SDD_{loc}$) and long-distance ($LDD_{loc}$) local seed dispersal events for \textit{Prunus mahaleb}.}
\end{figure}

%\section*{References}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
%%% Unnumbered Literature Cited
\bibliographystyle{bes}        %Compile with bes.bst style file
\bibliography{MS_BES}
